Analyzing the Impact of Social Attributes on Commit Integration Success

### 2017-03-28 Paper/MSRChallenge2017.tex
made the term data set consistent

\vspace{1ex}
\noindent\textbf{Metric.} We measure software contribution quality by the
build status of the
commit. We used a decision tree classifier~\cite{Quinlan86} to identify
important factors in predicting build success. Decision trees are a predictive 
model which uses binary decisions on the
attributes to predict a final class. They are well
established and interpretable. We use a J48 decision tree, which
has the ability to identify which attributes 
discriminate the various instances most clearly.
We first build a
decision tree classifier using only technical factors to 
predict build integration
success.  We then include the social factors, to 
identify the degree to which the social factors improve the model's predictive
power.
We evaluate the models with 10-fold cross-validation.

@@

We built a data set of
attributes consisting of both technical and social factors:

\vspace{1ex}
\noindent\textbf{Technical Factors.} We collected technical factors from the 
Mining Software Repositories (MSR) Challenge data set~\cite{msr17challenge}.
We extracted the following techical attributes as proxy to a commit's
complexity: 
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item Source churn, changed lines of code.
\item Files added by this commit.
\item Files modified by this commit.
\item Tests added by this commit.
\item Tests deleted by this commit.
\item Total source files in project at the time of the commit.
\item Number of source lines of code in the project at
	the time of the commit.
\item Travis status, or status of the Travis build integration.
\end{itemize}

The first five factors represent the changes caused by the commit.  The next two
factors cover the size of the project.  These factors represent the complexity
of the change, and the complexity of the project incorporating the change.

\vspace{1ex}
\noindent\textbf{Social Factors.} We created a web scraper to extract the social data
from developer's profiles.
We used the commit hash and the project name in the 
MSR Challenge~\cite{msr17challenge} dataset to locate the
commit web page, leveraging \textit{github.com}'s URL standard for projects and
commit hashes.
%  using the syntax
% below. 
% \newline \newline
% \textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
% \newline \newline
%
We scraped the commit author's
\textit{username} from each commit page to
collect the following from the user's public profile:
%(\textit{https://github.com/\textbf{UserName}}). 
%

@@

We also added one item from the MSR Challenge data set, indicating
whether the commit's developer is a core member  
of the associated project.
These social factors are all tied to a GitHub user's profile,
and are generally correlated with user time and engagement with the site.
Thus, we use them as a proxy for
experience.  If we were unable to find information about an author, we
assigned empty values to the commit.

@@

The challenge dataset~\cite{msr17challenge} consists
of commits that date as far back as 2011. The social data we 
gathered is accurate for January 2017. To mitigate the threat of
using future results to predict previous events,
we only include commits from the years 2014--2016.  Our intuition is
that these social attributes are unlikely to have significant recent changes.

We converted all Travis build statuses to two possible values:
\emph{successful} (passed in the challenge dataset) and \emph{unsuccessful}
(errored, failed, canceled
or started in the original data set). We also condensed all passing and
failing builds in a commit to a single pass or fail for each build status.  If
the commit contained both passing and failing build statuses, we treat it as
both a passing and failed build.

74.43\% of the builds in the reduced dataset were successful, and thus a naive
classifier that always guesses ``successful'' would succeed 74.43\% of the time.
To penalize misclassification for
both classes equally, we created a second \emph{balanced} dataset by randomly 
sub-sampling the successful builds until the two build status were
equal, through a process known as downsampling. We downsampled the entire
dataset and then trained decision tree classifiers on both.  We validated the
classifiers with cross-validation. Since the cross-validation is stratified, the
proportion of 50-50 is maintained through the training and testing folds for the
downsampled data. We present results for both
datasets in Section~\ref{sec:successClassification},  and only the full dataset for Sections~\ref{sec:core} and \ref{sec:followerSuccess}.

@@

\hline
\end{tabular}
\\
\center
	\caption{\label{resultsTable} Predictive power 
      decision trees with a non-downsampled dataset
	(2\textsuperscript{nd} and 3\textsuperscript{rd} column) and downsampled
	(4\textsuperscript{th} and 5\textsuperscript{th} column). 
	%The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}
	%Zack: the above sentence isn't true for the second row.
	}   
\end{table*}

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm.  Without downsampling, the 
decision tree shows an increase of 4.09\% \textit{correctly classified} instances when adding social data. On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, which outperformed the baseline of
guessing all builds passed (50.00\%).
 After including the social data
in the dataset, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.  These results show that
both technical and social data can be used to improve predictions of build 
integration success.

Note also that the Kappa statistic increases in both cases when
adding social data (100\% and 85\% respectively). The Kappa statistic accounts
for the accuracy of the model's predictive power. It is more
robust than \textit{correctly classified} instances alone, since it takes into account the
correctly guessed instances occurring by chance. The precision, recall and
F-measure are calculated over the successful and unsuccessful classes and then
averaged (standard procedure in Weka).  All these measures increase with social
data in the feature space, making the prediction more reliable. The mean 
absolute error decreases in both cases, further substantiating the importance 
of social factors
in predicting build success.

@@

The decision tree results suggested that core team members are
important to classifying build success, and thus we investigated the question
directly. 
We classified the build success rate for
core team members and non-core team members using the data set without 
downsampling.  We found that core team members
triggered 89.39\% (331827/371208) of the builds in the data set and 75.58\%
(250808/331827) of those builds
passed.  Non-core team members triggered 10.61\% (39381/371208) of the builds 
in the data set
and 64.70\% (25479/39381) of the builds passed.  The p-value from a Fisherexact
test 0.0000, which means that this difference is extremely unlikely to occur by
chance. These results show core team 
members are more
likely to have a successful build integration than non-core team members. We hypothesize the reason for this to be that core team members are more familiar with the code, therefore are more likely to have their builds pass.

@@

%\paragraph{Internal validity}
We use Weka, a well known machine-learning toolkit,
to analyze our data.
Like all tools, Weka may contain errors and/or bugs. 
Nonetheless, it is a well-known used, open source tool, and as such is likely
to be of relatively high quality, mitigating threats to internal validity. 
We have made our scripts, which analyzed the datasets and
obtained social data publicly available\footnote{https://goo.gl/CqTX44}, for
extension and reproducibility.
We mitigate the risk that our results fail to 
%\paragraph{External validity} 
generalize externally by studying a broad and rich 
 dataset~\cite{msr17challenge} comprising 1,359 popular projects.
Similarly, the social factors we study may not apply to closed-source
applications; however, open-source ecosystems are intellectually
valuable sources of study on their own. 
%
 Other concerns include the possibility that social factors
 have changed significantly in the past three years, that unsuccessful
 builds are due to
 incorrect test cases (rather than low quality contributions), and that
 information is lost in collapsing commits' multiple builds. We believe these
 are similarly mitigated by a large and varied dataset, but leave investigation
 of the effects of these decisions to future work.

@@

The most similar related work to this paper is a case study investigating how
social, technical, coordination, and prior-certification factors affect build
success for a project with 40 developers at IBM Toronto Labs~\cite{Hassan06}.  This work 
focused on one project and measured experience as the number of contributions to
the project. We focus on a wide range of open source projects and
measure experience through developer profile information.
Another study of the relationship between authors, files in project,
and team organization and build
success on an IBM project with 151 developers found that as the
number of authors or files in a build increases, the probability of build
success decreases~\cite{Kwan11}. A separate investigation of the
same project 
found that team structure affected the teams' build success rate~\cite{Wolf09}.
Others have established relationships between both social and technical attributes and 
different measures of success, such as pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse} or evaluation latency~\cite{Yu15}.
Tsay et al.~\cite{tsay14icse} 
show that the social connection between the developer initiating a
pull request and the project integrators is a key factor in predicting 
acceptance of a pull 
request. Unlike this previous study, we look at a broader aspect of 
social factors, and focus on predicting the Travis CI build success.
Gousios et al.~\cite{gousios14} found that the commit change location is
an important indicator to commit acceptance rate, and that including
test cases increases a
commit's acceptance rate~\cite{gousios15}.
Bettenburgh et 
al.~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, while Vasilescu et al.~\cite{vasilescu14} explore how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated into the automatic build process. Yu et al.~\cite{yu16} 
analyze the impact of CI usage in GitHub and its correlation with software 
quality. By contrast with their work, we focus on the relationship between independent social 
attributes and continuous integration success.

@@

We have analyzed how technical and social factors can be used to
predict high quality open source contributions, as measured by continuous
integration build success.  We have found that these factors provide
insight when predicting the success of a build with a technical factors leading
to a 14\% increase in predictive power, and another 12\% increase when
incorporating social factors into the prediction.

Overall, we conclude that social attributes proxying experience, such as whether
the developer is a core project contributor,
noticeably influence build success prediction.  While the increase was noticeable 
with the full data set, there was a much greater increase in prediction
accuracy after downsampling. 
%These results suggest that social factors
%are significant predictors when
%evaluating the predictive power of build commit success.
%
In particular, core team members are
more likely to have the build integration pass than non-core team members.  This
result provides evidence that non-core team member commits may not be as
scrutinized as core team member commits before the integration build tests,
although further study is required to draw definitive
conclusions.
%
Likewise, developers with more followers are more
likely to pass the build, although followers are not a perfect indicator.
Developers with 1000+ followers were more than a 10\% more succesful in their
builds as compared to
developers with no followers.  This suggests that open source developers with
high quality commits may be more
likely to have a large number of followers.

>>>>

\vspace{1ex}
\noindent\textbf{Metric.} We measured software contribution quality by the
build status of the
commit. We used a decision tree classifier~\cite{Quinlan86} to predict build 
success. Decision trees 
use binary decisions on the
attributes to predict a final class. They are well
established and interpretable. We used a WEKA J48 decision tree for readability.
We first built a
decision tree classifier using only technical factors to 
predict build integration
success.  We then included the social factors, to 
identify the degree to which social factors improve the model's predictive
power.
We evaluated the models with 10-fold cross-validation.

@@

We built a data set consisting of both technical and social factors:

\vspace{1ex}
\noindent\textbf{Technical Factors.} We collected technical factors from the 
Mining Software Repositories (MSR) Challenge data set~\cite{msr17challenge}.
We extracted the following technical attributes to proxy a commit's
complexity: 
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item Source churn, changed lines of code.
\item Files added by this commit.
\item Files modified by this commit.
\item Tests added by this commit.
\item Tests deleted by this commit.
\item Total source files in project at the time of the commit.
\item Number of source lines of code in the project at
	the time of the commit.
\item Travis status, or status of the Travis build integration.
\end{itemize}

The first five factors represent the changes caused by the commit.  The next two
factors cover the size of the project.  These factors represent the commit's complexity
, and the complexity of the project incorporating the commit.

\vspace{1ex}
\noindent\textbf{Social Factors.} We created a web scraper to extract social data
from developer's profiles.
We used the commit hash and the project name in the 
MSR Challenge~\cite{msr17challenge} data set to locate the
commit web page, leveraging \textit{github.com}'s URL standard for projects and
commit hashes.
%  using the syntax
% below. 
% \newline \newline
% \textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
% \newline \newline
%
We scraped the commit author's
\textit{username} from each commit page to
collect the following from the user's public profile:
%(\textit{https://github.com/\textbf{UserName}}). 
%

@@

We also added one item from the MSR Challenge data set, indicating
whether the commit's developer is a core member  
of the associated project.
These social factors are all tied to a GitHub user's profile,
and are generally correlated with users' open source engagement.
Thus, we use them as a proxy for
experience.  If we were unable to find information about an author, we
assigned empty values to the commit.

@@

The challenge data set~\cite{msr17challenge} consists
of commits that date as far back as 2011. The social data we 
gathered is accurate for January 2017. To mitigate the threat of
using future results to predict previous events,
we only include commits from the years 2014--2016.  Our intuition is
that these social attributes are unlikely to have significant recent changes.

We converted all Travis build statuses to two possible values:
\emph{successful} (passed in the challenge data set) and \emph{unsuccessful}
(errored, failed, canceled
or started in the original data set). We also condensed all passing and
failing builds in a commit to a single pass or fail for each build status.  If
the commit contained both passing and failing build statuses, we treat it as
both a passing and failed build.

74.43\% of the builds in the reduced data set were successful, and thus a naive
classifier that always guesses ``successful'' would succeed 74.43\% of the time.
To penalize misclassification for
both classes equally, we created a second \emph{balanced} data set by randomly 
sub-sampling the successful builds until the two build status were
equal, through a process known as downsampling. We downsampled the entire
data set and then trained decision tree classifiers on both.  We validated the
classifiers with cross-validation. Since the cross-validation is stratified, the
proportion of 50-50 is maintained through the training and testing folds for the
downsampled data. We present results for both
data sets in Section~\ref{sec:successClassification},  and only the full data 
set in Sections~\ref{sec:core} and \ref{sec:followerSuccess}.

@@

\hline
\end{tabular}
\\
\center
	\caption{\label{resultsTable} Predictive power 
      decision trees with a non-downsampled data set
	(2\textsuperscript{nd} and 3\textsuperscript{rd} column) and downsampled
	(4\textsuperscript{th} and 5\textsuperscript{th} column). 
	%The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}
	%Zack: the above sentence isn't true for the second row.
	}   
\end{table*}

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm.  Without downsampling, the 
decision tree\footnote{decisions trees and datasets: } shows an increase of 4.09\% \textit{correctly classified} instances when adding social data. On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, which outperformed the baseline of
guessing all builds passed (50.00\%).
After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.  These results show that
both technical and social data can be used to improve predictions of build 
integration success.

The precision, recall and
F-measure are calculated over the successful and unsuccessful classes and then
averaged (standard procedure in Weka).  All these measures, along with the Kappa
statistic, increase with social features, which provide evidence that social 
factors improve build success predictions. The mean 
absolute error decreases in both cases, further substantiating the importance 
of social factors
in predicting build success.

@@

The decision tree results suggested that core team members are
important to classifying build success, and thus we investigated the question
directly. 
We classified the build success rate for
core team members and non-core team members using the data set without 
downsampling.  We found that core team members
triggered 89.39\% (331827/371208) of the builds in the data set and 75.58\%
(250808/331827) of those builds
passed.  Non-core team members triggered 10.61\% (39381/371208) of the builds 
in the data set
and 64.70\% (25479/39381) of the builds passed.  The p-value from a Fisher exact
test 0.0000, which means that this difference is extremely unlikely to occur by
chance. These results show core team 
members are more
likely to have a successful build integration than non-core team members. We
hypothesize this occurs because core team members are more familiar with the 
code, and therefore more likely to have their builds pass.

@@

%\paragraph{Internal validity}
We use Weka, a well known machine-learning toolkit,
to analyze our data.
Like all tools, Weka may contain bugs. 
Nonetheless, it is a community tested open source tool, which mitigates threats to 
internal validity. 
We have made our scripts, which analyzed the data sets and
obtained social data publicly available\footnote{https://goo.gl/CqTX44}, for
extension and reproducibility.
We mitigate the risk that our results fail to 
%\paragraph{External validity} 
generalize externally by studying a broad and rich 
 data set~\cite{msr17challenge} comprising 1,359 popular projects.
Similarly, the social factors we study may not apply to closed-source
applications; however, open-source ecosystems are intellectually
valuable sources of study on their own. 
%
 Other concerns include the possibility that social factors
 have changed significantly in the past three years, that unsuccessful
 builds are due to
 incorrect test cases (rather than low quality contributions), and that
 information is lost in collapsing commits' multiple builds. We believe these
 are similarly mitigated by a large and varied data set, but leave investigation
 of the effects of these decisions to future work.

@@

The most similar related work to this paper is a case study investigating how
social and technical factors affect build
success on a project with 40 developers~\cite{Hassan06}.
In contrast to this work, we focus on a wide range of open source projects and
measure experience through developer profile information.
Other work has investigated the effects of various factors, such as team
organization, on predicting build success in a single
project~\cite{Kwan11,Wolf09}.
Previous investigations have established relationships between both social and 
technical 
attributes and 
different measures of success, such as pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse} or evaluation latency~\cite{Yu15}.
Gousios et al.~\cite{gousios14} found that the commit change location is
an important indicator to commit acceptance rate, and that including
test cases increases a
commit's acceptance rate~\cite{gousios15}.
Bettenburgh et 
al.~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, while Vasilescu et al.~\cite{vasilescu14} explore how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated into the automatic build process. Yu et al.~\cite{yu16} 
analyze the impact of CI usage in GitHub and its correlation with software 
quality. By contrast with their work, we focus on the relationship between social 
attributes and continuous integration success.

@@

We have analyzed how technical and social factors can be used to
predict high quality open source contributions, as measured by continuous
integration build success.  We have found that these factors can
insight when predicting the success of a build; technical factors led
to a 14\% increase in predictive power, and adding social factors led to another
12\% increase.

Overall, we conclude that social attributes proxying experience, such as whether
the developer is a core project contributor,
noticeably influence build success prediction, especially on the downsampled
data. 
%These results suggest that social factors
%are significant predictors when
%evaluating the predictive power of build commit success.
%
In particular, core team members are
more likely to have the build integration pass than non-core team members.  This
result provides evidence that non-core team member commits may not be
scrutinized as closely as core team member commits before the integration build 
tests,
although further study is required to draw definitive
conclusions.
%
Likewise, developers with more followers are more
likely to pass the build.
Developers with 1000+ followers were more than 10\% more succesful in their
builds compared to
developers with no followers.  This suggests that open source developers with
high quality commits may be more
likely to have a large number of followers.
